### 2025-08-30 getting-started.qmd
20250830 - merge explanation and visual

>>>>

If we have two data frames that we want to combine into one we would perform what is called a merge.
Let's say we want to examine how players' fantasy performance differs by age.
Assume we have players' age in one data frame and players' fantasy points in a separate data frame.
To examine how age and fantasy performance is related, we will need to have both variables in the same data frame.
Thus, to examine age in relation to fantasy performance, we would need to merge both data frames so that the variables age and fantasy points are in the same data frame.
However, to merge successfully, we need to indicate how the rows in the first data frame can be connected to the rows in the second data frame.
For instance, if the player's ID is a variable in both data frames, we could use the player's ID as a "key" to indicate how to merge the rows from both data frames.
The merged object, then, would retain the keys (i.e., player ID) and the variables from each object (i.e., age and fantasy points).
An example of merge the two data frames in this example is depicted in @fig-merge.

::: {#fig-merge}
![](images/merge.png){width=100% fig-alt="Example of Merging Two Data Frames."}

"Example of Merging Two Data Frames."
:::

Here is a code example:

```{r}
dataframeA <- data.frame(
  id = c(1, 2, 3),
  age = c(24, 26, 30)
)

dataframeB <- data.frame(
  id = c(1, 2, 3),
  fantasyPoints = c(200, 300, 250)
)

mergedDataframe <- dplyr::full_join(
  dataframeA,
  dataframeB,
  by = "id"
)

dataframeA

dataframeB

mergedDataframe
```
